---
title: "R Workshop"
author: "Muji Chughtai"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

# Base Functions in R

```{r}
17 + 18
5 * 4
100/15
4^3

# Statistics (percentage below a z-score of -1)
pnorm(q = -1, mean = 0, sd = 1, lower.tail = TRUE)

## Percentage below a z-score of -3, -2, -1, 0, 1, 2, and 3
pnorm(q = c(-3,-2,-1,0,1,2,3), mean = 0, sd = 1, lower.tail = TRUE)
```

```{r}
# Using variables
a <- 3
b <- 4
c <- sqrt(a^2 + b^2); c

a <- c(3, 6, 9, 12, 15)
b <- c(4, 8, 12, 16, 20)
perfect.hyps <- sqrt(a^2 + b^2)
perfect.hyps
```

```{r}
# Using variables in R's base functions to compare scores on different standardized tests

## ACT
pop.mu = 20.8
pop.sd = 5.8

my.score <- 30

pnorm(q = my.score, mean = pop.mu, sd = pop.sd)
### Note that you do not need to include the parameter names in the function call
pnorm(my.score, pop.mu, pop.sd)

## GRE
pop.mu <- 150
pop.sd <- 8.6

my.scores <- c(158, 160, 165)

pnorm(my.scores, pop.mu, pop.sd)
```

# Load Required Packages and Data

```{r}
library(tidyverse)
RawData <- read.csv("RWorkshop.csv")
```

# Data Cleaning
## Remove Unnecessary Rows

```{r}
# Removing rows 1 through 2 - Note the comma after the parentheses
CleanerData <- RawData[-c(1:2),]
```

## Remove Unnecessary Columns
```{r}
# Retaining only the columns that we need
CleanData <- CleanerData %>%
  select(FirstName, LastName, Region, Birthday.1_1, Birthday.2_1, Birthday.3_1, AgeYears, HeightInches)
```

## Calculating Length of Names
```{r}
# Calculating the number of characters (letters) in each name
Name.Data <- CleanData %>%
  mutate(FirstNameLength = nchar(FirstName),
         LastNameLength = nchar(LastName))
```

## Simplifying `Region`
```{r}
# Removing states from the region values
Region.Data <- Name.Data %>%
  mutate(Region = case_when(startsWith(Region, "Northeast") ~ "Northeast",
                            startsWith(Region, "Midwest") ~ "Midwest",
                            startsWith(Region, "Southeast") ~ "Southeast",
                            startsWith(Region, "Southwest") ~ "Southwest",
                            startsWith(Region, "West") ~ "West",
                            startsWith(Region, "International") ~ "International"))
```

## Simplifying `Birthday`
```{r}
# Combining the three birthday variables into a single one and converting to R date format; removing the original (now unnecessary) birthday variables
Bday.Data <- Region.Data %>%
  mutate(Birthday = paste0(Birthday.1_1, " ", Birthday.2_1, ", ", Birthday.3_1)) %>%
  mutate(Birthday = as.Date(Birthday, format = "%B %d, %Y")
         ) %>%
  select(-Birthday.1_1, -Birthday.2_1, -Birthday.3_1)
```

## Calculating Age in Days and Height in CMs
```{r}
# Subtracting current date from birthday to calculate age in days as a numeric variable; ensuring height is numeric and converting height from feet to inches for individuals who accidentally reported their height as feet (since it is less than 7); converting height in inches to height in CMs; removing individuals who have an NA for height in cms or age in days (our main DVs)
FinalData <- Bday.Data %>%
  mutate(AgeDays = as.numeric(difftime(Sys.Date(), Birthday, units = "days")),
         HeightInches = as.numeric(HeightInches),
         HeightInches = ifelse(HeightInches < 7, HeightInches*12, HeightInches),
         AgeYears = as.numeric(AgeYears),
         HeightCMs = HeightInches*2.54) %>%
  filter(!is.na(HeightCMs) | !is.na(AgeYears))
```

# Data Anlysis and Visualization

## Comparing Height in CM and Height in Inches
```{r}
# Correlation
cor(FinalData$HeightCMs, FinalData$HeightInches, method = "pearson")

# Linear Regression
model1 <- glm(HeightCMs ~ HeightInches, data = FinalData)
summary(model1)
```

### Visualizing this Relationship
```{r, message = FALSE}
# Density of predictor variable (IV)
FinalData %>%
  ggplot(aes(x = HeightInches)) +
  geom_density(fill = "#BDD7E7") + theme_minimal() +
  labs(title = "Height in Inches of Summer 2024 Interns",
         x = "Height in Inches")

# Linear relationship of IV and DV
FinalData %>%
  ggplot(aes(x = HeightInches, y = HeightCMs)) +
  geom_point() +
  geom_smooth(method = "lm")+
  labs(title = "The Perfect Linear Relationship between Height in CMs and Height in Inches",
       subtitle = "\U03B2 = 2.54",
       x = "Height in Inches",
       y = "Height in CMs") +
  theme_minimal()
```

## Comparing Age in Days and Age in Years
```{r}
# Correlation
cor(FinalData$AgeDays, FinalData$AgeYears, method = "pearson")

# Linear Regression
model2 <- glm(AgeDays ~ AgeYears, data = FinalData)
summary(model2)
```

### Visualizing this Relationship
```{r, message = FALSE}
# Density of predictor variable (IV)
FinalData %>%
  ggplot(aes(x = AgeYears)) +
  geom_density(fill = "#BDD7E7") + theme_minimal() +
  labs(title = "Age in Years of Summer 2024 Interns",
         x = "Age in Years")

# Linear relationship of IV and DV
FinalData %>%
  ggplot(aes(x = AgeYears, y = AgeDays)) +
  geom_point() +
  geom_smooth(method = "lm")+
  labs(title = "Age in Years by Age in days",
       subtitle = "\U03B2 ≈ 365",
       x = "Age in Years",
       y = "Age in Days") +
  theme_minimal()
```

## Comparing Length of First Names and Last Names
```{r}
# Correlation
cor(FinalData$FirstNameLength, FinalData$LastNameLength, method = "pearson")

# Linear Regression
model2 <- glm(LastNameLength ~ FirstNameLength, data = FinalData)
summary(model2)
```

### Visualizing this Relationship
```{r, message = FALSE}
# Density of predictor variable (IV)
FinalData %>%
  ggplot(aes(x = FirstNameLength)) +
  geom_density(fill = "#BDD7E7") + theme_minimal() +
  labs(title = "Length of First Names of Summer 2024 Interns",
         x = "Length of First Names")

# Linear relationship of IV and DV; note that we use geom_jitter instead of geom_point to see multiple points that have the same (x,y) coordinates. For example, there are multiple individuals with 6 character first names and 4 character last names, and we want to be able to see them all and not just one point at (6,4)
FinalData %>%
  ggplot(aes(x = FirstNameLength, y = LastNameLength)) +
  geom_jitter(height = 0.1, width = 0.1) + 
  geom_smooth(method = "lm")+
  labs(title = "Last Name Length by First Name Length",
       subtitle = "\U03B2 ≈ 0",
       x = "First Name Length",
       y = "Last Name Length") +
  theme_minimal()
```

## Height by Region
```{r}
# Table
height.by.region <- FinalData %>%
  group_by(Region) %>%
  summarise(Mean.Height = mean(HeightCMs),
            N = n()) %>%
  arrange(desc(Mean.Height))

knitr::kable(height.by.region)
```

```{r}
# Linear Model (ANOVA)
## Reordering the levels in order of descending mean
FinalData$Region <- factor(FinalData$Region, levels = height.by.region$Region)

model3 <- glm(HeightCMs ~ Region, data = FinalData)
summary(model3)
```

### Visualizing this Relationship

```{r, message = FALSE}
FinalData %>%
  ggplot(aes(x = Region, y = HeightCMs, fill = Region)) +
  scale_fill_brewer(palette = "Spectral") +
  geom_boxplot() +
  labs(title = "Height in CMs by Region",
       x = "Region",
       y = "Height in Cms") +
  theme_minimal() +
  theme(legend.position = "none")
```